\chapter{Experiments design and results}
\label{chap:expe}
\textit{In this Chapter, we expose all details about the evaluation process of \posl{}, \ie all experiments performed. For each benchmark, we explain strategies used in the evaluation process and what are the experiment environments before exposing a complete analysis of the obtained results.}
\vfill
\minitoc
\newpage

In this chapter, we illustrate and analyze the versatility of \posl{} by studying different ways to solve constraint problems based on local search meta-heuristics. 
We have chosen the \sgp, the \nqp, the \carrp{} and the \grp{} as benchmarks since they are challenging yet differently structured problems. 
\sgp{} has the structure of tournament problems, where the scheduling of matches between players along a given period of time. The constraints, related to how many times a player can participate in a match the same week, make the problem more complex as the number of weeks increases. \nq{} and \carr{} are similarly modeled in these experiments, since they are represented as permutation problems. However, they have a very interesting characteristic which differentiate them from each other: from certain order on, the number of solutions with respect to the order increases for the case of \nqp{}, and for decrease drastically for \carrp{}. \grp{} was chosen for two main reasons. \new{Its solution representation is very different from the other studied problems, and because during the search process, \posl{} describes a different behavior: it performs many restarts.}
%The first one is because unlike benchmarks mentioned before, where their solution representation's are permutations of $s = (0, 1, \dots, n)$ (or blocks of permutations for the case of \sg), the solution representation of \grp{} is just an ordered set of integers. }

%In this chapter we present formally each benchmark, we explain the structure of \posl's solvers that we have generated for experiments and present a detailed analysis of obtained results.

First results using \posl{} to solve constraint problems were published in \cite{Reyes-amaro} where it was used to solve the \sgp{} and to study some communication strategies. It was the first version of \posl{}, and it was able to solve relatively easy instances only. However, results suggested that the communication can play an important role if we are able to find the proper \commstr{}, and they encourage us to go even further on this direction. In this first version were implemented simple communication mechanisms only (communication \oneTone{} and \oneTn{}, but only in one direction.)

In the second and current version of \posl{}\footnote{\posl{} source code is available on GitHub: \href{https://github.com/alejandro-reyesamaro/POSL}{https://github.com/alejandro-reyesamaro/POSL}}, other \commopers{} were incorporated, and those already existent were improved, in order to be able to build more sophisticated \commstrs. With this version, we started deeper investigations with more advanced experimentations, \new{which are presented in this chapter.} 

\section{Methodology}

Some terms are necessary to be defined for simplification sake: the \textit{sequential environment} and the \textit{parallel environment} are the description of the computation resources used for experimentation. Experiments were performed on an Intel\R{} Xeon\TM{} E5-2680 v2, 10$\times$4 cores, 2.80GHz. This server is called {\sc Curiosiphi} and is located at the  \textit{Laboratoire des Sciences du Numérique de Nantes (LS2N)} of the Université de Nantes. %The other term that needs to be introduced is the \textit{statistical summary}, which is the obtained information from the runs, used to analyze the solution process and draw conclusions.

\begin{definition}\label{def:seq_envir}
We say that we launch an experiment using a \textbf{sequential environment} if we execute a \soset{} into a single process of {\sc Curiosiphi}.
\end{definition}

\begin{definition}\label{def:par_envir}
We say that we launch an experiment using a \textbf{parallel environment} if we execute a \soset{} in parallel (multi-walk) using the maximum of available processes in {\sc Curiosiphi}.
\end{definition}

%\begin{definition}\label{def:summ_stat}
%During the experimentation process, we say that we are performing a \textbf{statistical summary} if  
%\end{inparaenum} 
%\end{definition}
%.

With the aim of being as exhaustive as possible in the experimentation process, a methodology based on four stages is proposed:

\begin{enumerate}
\item \underline{\textbf{Algorithm selection}} In this stage some experiments are launched, using the sequential environment, to ensure choosing the right \oms, and the right design of the \as. The following statistical analysis is performed: A set of 30 runs for each setup are performed, and used to \begin{inparaenum}[a)]
\item build box-plot diagrams and bars graphs with some additional information about winner solvers, presented in Appendixes~\ref{app:sgp}, \ref{app:nqp}, \ref{app:cap} and \ref{app:grp}; \item compute means and standard deviation for run-times and iterations, showed in tables, in columns labeled {\bf T} (run-time in seconds), {\bf It.} (number of iterations), {\bf T(sd)} and {\bf It.(sd)} (their respective standard deviations). In some tables, the column labeled \textbf{\% success} indicates the percentage of \sosets{} finding a solution before reaching a time--out of 5 minutes (imperative when dealing with meta-heuristics). \end{inparaenum}

\item \underline{\textbf{Algorithm evaluation in the parallel environment}} The selected algorithm is launched using the parallel environment. It is performed a similar statistical analysis to the one described in the previews stage, and results are compared.\label{stage_expe_2}

\item \underline{\textbf{Communication strategies selection}} After a detailed study of the search process and the behavior of the designed \sosets, some changes in the \soset{} are proposed in order to design a \commstr:
\begin{itemize}
\item replacing some \oms{} for others based on the originals, but with some modifications according to the new demands of the proposed \commstr;
\item adding some \opchs{} depending on the information that we intend to share;
\item a new \as{} is coded, whose modifications are the strictly necessary to incorporate \opchs;
\item the structure of the communication is designed in order to chose the right \commopers.
\end{itemize}

\item \underline{\textbf{Communication strategy evaluation}} The designed \commstr{} is launched using the parallel environment, and a statistical analysis is performed. Communication strategies are compared each others based on obtained results in order to select the right one. These results are also compared to those obtained during Stage~\ref{stage_expe_2}, to be able to draw conclusions about the success of the cooperative approach.
\end{enumerate}

It is important to point out that \posl{} is designed neither to obtain the best results in terms of performance, nor to outperform the state-of-the-art solutions, but to give the possibility of rapidly prototyping and studying different cooperative or non cooperative search strategies.

%All benchmarks were coded using the \posl{} low-level framework in C++.

%\section{Solving the \sgp}\label{sec:golfers} % Studying a dynamic configuration exchange strategy with
\section{A dynamic configuration exchange strategy (\sg)}
\input{results/res_golfers}

\section{A cyclic \comstr{} (\nq)}\label{sec:nqueens} % Studying a cyclic \comstr{} with
\input{results/res_nqueens}

\section{A simple \comstr{} (\carr)}\label{sec:costas} % Studying a simple \comstr{} with
\input{results/res_costas}

\section{A local minima evasion strategy (\gr)}\label{sec:golomb} % Studying a local minima evasion strategy with Golomb Ruler
\input{results/res_golomb}

\section{Summarizing}

In this chapter various \CSPs{} have been chosen to \begin{inparaenum}[1)] \item evaluate the \posl{} behavior solving these kind of problems, and \item study different solution strategies, specially communication strategies. \end{inparaenum} To this end, benchmarks with different characteristics have been selected to help getting a wide view of the \posl{} behavior.

In the solution process of \sgp{}, we studied an exploitation-oriented \commstr, in which the current configuration is communicated to ask other solvers for help to concentrate the effort in a more promising area. Results show that this \commstr{} can provide some gain in terms of runtime. It was also presented results showing the success of a cost descending acceleration \commstr, exchanging the current configuration between two solvers with different characteristics. Some other unsuccessful \commstrs{} were studied, showing that the sub-division of the effort by weeks does not work well. Table~\ref{tab:golfers_summarize} summarizes the results obtained solving \SGP. 

\begin{table}[h]
\captionsetup{belowskip=6pt,aboveskip=6pt}
\centering 
\renewcommand{\arraystretch}{1}
\begin{tabular}{p{1.7cm}|R{0.8cm}R{1cm}|R{0.8cm}R{1cm}|R{0.8cm}R{1cm}}
	\hline 	
	\multirow{2}{*}{\centering {\bf Instance}} & \multicolumn{2}{c|}{\bf Sequential} & \multicolumn{2}{c|}{\bf Parallel} & \multicolumn{2}{c}{\bf Cooperation}\\
	\cline{2-7}
	& T & It. & T & It. & T & It. \\
	\hline
	5--3--7 & 1.25 & 2,907 & 0.23 & 142 & \good{0.08} & 139 \\
	8--4--7 & 0.60 & 338 & 0.28 & 93 & \good{0.14} & 100 \\
	9--4--8 & 1.04 & 346 & 0.60 & 139 & \good{0.36} & 144 \\
	\hline
\end{tabular}
\caption{Summarizing results for \SGP}
\label{tab:golfers_summarize}
\end{table}

It was showed that simple communication strategies as they were applied to solve \sgp{} do not improve enough the results without communication for the \nqp{}. %In this problem, the number of solution with respect to the order $N$ increase exponentially, then higher order instances are "easer" to solve using local search. For that reason, the communication can not provide a lot of gain. 
However, a deep study of the \posl's behavior during the search process allowed to design a communication strategy able to improve obtained results using non-communicating strategies for small instances. Table~\ref{tab:nqueens_summarize} summarizes the obtained results solving \NQP.

\begin{table}[h]
\captionsetup{belowskip=6pt,aboveskip=6pt}
\centering 
\renewcommand{\arraystretch}{1}
\begin{tabular}{p{1.7cm}|R{0.8cm}R{1cm}|R{0.8cm}R{1cm}|R{0.8cm}R{1cm}}
	\hline 	
	\multirow{2}{*}{\centering {\bf Instance}} & \multicolumn{2}{c|}{\bf Sequential} & \multicolumn{2}{c|}{\bf Parallel} & \multicolumn{2}{c}{\bf Cooperation}\\
	\cline{2-7}
	& T & It. & T & It. & T & It. \\
	\hline
	250 & 0.29 & 8,898 & 0.19 & 4,139 & \good{0.09} & 1,169 \\
	500 & 0.35 & 4,203 & 0.24 & 2,675 & \good{0.14} & 864 \\
	1000 & 0.35 & 2,766 & 0.30 & 2,102 & \good{0.21} & 807 \\
	3000 & 1.50 & 2,191 & 1.33 & 2,168 & \good{1.02} & 1,613 \\
	6000 & 4.71 & 3,339 & 4.57 & 3,323 & \good{4.24} & 2,537 \\
	\hline
\end{tabular}
\caption{Summarizing results for \NQP}
\label{tab:nqueens_summarize}
\end{table}

The \carrp{} is a very complicated constrained problem, and very sensitive to the methods to solve it. For that reason we used some ideas from already existing algorithms. However, thanks to some studies of different \commstrs, based on the communication of the current configuration at different times (places) in the algorithm, it was possible to find a communication strategy to improve the performance. Table~\ref{tab:costas_summarize} summarizes the results for solving \CARRP. 

\begin{table}[h]
\captionsetup{belowskip=6pt,aboveskip=6pt}
\centering
\renewcommand{\arraystretch}{1}
%\resizebox{0.9\columnwidth}{!}{%
\begin{tabular}{p{3.5cm}|R{1.2cm}R{1.7cm}R{1.7cm}}
	\hline
	{\bf STRATEGY} & T & It. & \% success\\
	\hline
	%\hline
	Sequential  & 132.73 & 2,332,088 & 40.00\\
	Parallel & 25.51 & 231,262 & 100.00\\
	Cooperative Strategy & \good{10.83} & 79,551 & 100.00\\
	\hline
\end{tabular}
%}
\caption{Summarizing results for \CARRP{} 19}
\label{tab:costas_summarize}
\end{table}

During the solution process of the \grp{}, \posl{} needs to perform many restarts. For that reason, this problem was chosen to study a different (and innovative up to our knowledge) communication strategy in which the communicated information is a potential local minimum to be avoided. This new communication strategy showed to be effective to solve this kind of problems. Table~\ref{tab:golomb_summarize} summarizes the results for solving \GRP.

\begin{table}[h]
\captionsetup{belowskip=6pt,aboveskip=6pt}
\centering 
\renewcommand{\arraystretch}{1}
\resizebox{\columnwidth}{!}{
\begin{tabular}{p{1.5cm}|R{1.1cm}R{1.3cm}R{0.6cm}R{1.7cm}|R{0.9cm}R{1.3cm}R{0.6cm}|R{1.1cm}R{1.3cm}R{0.6cm}}
\hline 	
{\bf Instance} & \multicolumn{4}{c|}{\textbf{Sequential}} & \multicolumn{3}{c|}{\textbf{Parallel}} & \multicolumn{3}{c}{\textbf{Cooperation}}\\
\cline{2-11}
& T & It. & R & \% success & T & It. & R & T & It. & R\\
\hline
8--34 & 0.66 & 10,745 & 53 & 100.00 & 0.43 & 349 & 1 & \good{0.43} & 283 & 1\\
10--55 & 67.89 & 446,913 & 297 & 88.00 & 4.92 & 20,504 & 13 & \good{3.16} & 12,605 & 8\\
11--72 & 117.49 & 382,617 & 127 & 30.00 & 85.02 & 155,251 & 51 & \good{60.35} & 110,311 & 36\\
\hline
\end{tabular}
}
\caption{Summarizing results for \GRP{}}
\label{tab:golomb_summarize}
\end{table}

In all cases, using the operator-based language provided by \posl{}, it was possible to test many different strategies (communicating and non-communicating) fast and easily. Whereas creating solvers implementing different solution strategies can be complex and tedious, \posl{} gives the possibility to make communicating and non-communicating solver prototypes and to evaluate them with few efforts. In this chapter we showed that a good selection and management of inter-solvers communication can %largely 
help to the search process, working with complex constrained problems.

\subsection{Strategies overview}
\input{results/res_extra_v2}