\thispagestyle{empty}
\newgeometry{left=3.2cm,right=2.3cm,top=2.8cm,bottom=2.8cm} % Change margins locally
\hfill\Huge\textsc{List of Papers}\\
\normalsize
\noindent\rule[2pt]{\textwidth}{0.8pt}\\
%\huge{\textsc{List of Papers\\}}\normalsize

\begin{enumerate}[I]
  \item Preus, S. and Wilhelmsson, L. M.\\
  \textit{Advances in Quantitative FRET-Based Methods for Studying Nucleic Acids}\\
  ChemBioChem (\textbf{2012}), 13, 1990-2001
  \item Preus, S., Börjesson, K., Kilså, K., Albinsson, B., and Wilhelmsson, L. M.\\
  \textit{Characterization of Nucleobase Analogue FRET Acceptor tC$_{nitro}$\\}
  J. Phys. Chem. B (\textbf{2010}), 114, 1050-1056
  \item Preus, S., Kilså, K., Miannay, F. A., Albinsson, B., and Wilhelmsson, L. M.\\
  \textit{FRETmatrix: a general methodology for the simulation and analysis of FRET in nucleic acids}\\
  Nucleic Acids Res. (\textbf{2012}), doi: 10.1093/nar/gks856
  \item Burns, J. R., Preus, S., Singleton, D. G., and Stulz, E.\\
  \textit{A DNA based five-state switch with programmed reversibility}\\
  Chem. Commun. (\textbf{2012}), 48, 11088-11090
  \item Preus, S., Kilså, K., Wilhelmsson, L. M., and Albinsson, B.\\
  \textit{Photophysical and structural properties of the fluorescent nucleobase analogues of the tricyclic cytosine (tC) family}
  Phys. Chem. Chem. Phys. (\textbf{2010}), 12, 8881-8892
  \item Dierckx, A., Miannay, F. A., Gaied, N. B., Preus, S., Björck, M., Brown, T., and Wilhelmsson, L. M.\\
  \textit{Quadracyclic Adenine: A Non-Perturbing Fluorescent Adenine Analogue\\}
  Chem. Eur. J. (\textbf{2012}), 18, 5987-5997
\end{enumerate}

%\vspace{2 cm}
\vfill
\Large\textsc{Contribution report}\normalsize

Papers I, II, III and V are mainly the work of the author. Paper IV is the main work of J. Burns and E. Stulz where the author contributed with simulations, data analysis and discussion of results. Paper VI is the main work of A. Dierckx and F. A. Miannay where the author contributed with design, calculations and discussion of results.

\newpage
\restoregeometry
\newgeometry{left=3cm,right=2cm,top=2.5cm,bottom=0.9cm} % Change margins locally
Papers not included in the thesis:

\begin{itemize}
  \item Preus, S., Jønck, S., Pittelkow, M., Monhaphol, T., Albinsson, B., and Wilhelmsson, L. M.\\
  \textit{The photoproduct of fluorescent nucleobase analogue tC in the presence of O$_2$}\\
  In preparation. (\textbf{2012})
  \item Preus, S., Börjesson, K., McPhee, S.A., Lilley, D.M.J., and Wilhelmsson, L.M.\\
  \textit{tC$^O$: a highly fluorescent RNA base analogue}\\
  In preparation. (\textbf{2012})
\end{itemize}
\vspace{1cm}
Papers published prior PhD:
\begin{itemize}
  \item Börjesson, K., Preus, S., El-Sagheer, A. H., Brown, T., Albinsson, B., and Wilhelmsson, L. M.\\
  \textit{Nucleic Acid Base Analog FRET-Pair Facilitating Detailed Structural Measurements in Nucleic Acid Containing Systems}\\
  J. Am. Chem. Soc. (\textbf{2009}), 131, 4288-4293
  \item Shanks, D., Preus, S., Qvortrup, K., Hassenkam, T., Nielsen, M. B., and Kilså, K.\\
  \textit{Excitation energy transfer in novel acetylenic perylene diimide scaffolds}\\
  New J. Chem. (\textbf{2009}), 33, 507-516
\end{itemize}

\vspace{2 cm}
 Software published during PhD (\url{www.fluortools.com}):

\begin{itemize}
  \item \textit{FRETmatrix:} Simulation and analysis of FRET (published with Paper III)
  \item \textit{a}|\textit{e:} UV-Vis absorption and emission spectra analysis software
  \item \textit{FluorFit:} Analysis of time-resolved fluorescence decays
  \item \textit{AniFit:} Analysis of time-resolved fluorescence anisotropy
\end{itemize}
\thispagestyle{empty}
\restoregeometry
%\end{vcentrepage}
